Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2018-12-03 Manuscript/manuscript.Rmd

---
title: "Manuscript title"
#csl: styles.ref/genomebiology.csl
output: 
  word_document:
    #reference_docx: C:/Users/Spiro Stilianoudakis/Documents/TAD_predictive_modeling/Manuscript/styles.doc/NIH_grant_style.docx
bibliography: C:/Users/Spiro Stilianoudakis/Documents/TAD_predictive_modeling/Manuscript/References/references.bib
---

@@

**Methods.** We propose a novel data pre-processing pipeline that addresses those shortcomings. A number of classifier performance metrics were assessed, including the F1 measure and Matthew Correlation Coefficient (MCC), and area under the ROC curve (AUROC).

**Results.** Data preprocessing (log2-transformation and standardization) improves the performance of classification algorithms and allows for the ability to more accurately predict which genomic features are most associated with TAD boundaries.

**Conclusions.** Current methods used to model the epigenomic features associated with TAD boundaries are insufficiently robust to handle properties of genomic data. Models applied to unprocessed data can have poor predictive performances. Focusing solely on typical performance assessment metrics, such as AUROC, can mask poor performance of the models; thus, the use of more balanced metrics, such as F1 and MCC, is warranted. Our model results in better performances and more accurate identification of important features associated with the formation of TAD boundaries.

>>>>

---
title: "Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings"
csl: C:/Users/Spiro Stilianoudakis/Documents/TAD_predictive_modeling/Manuscript/styles.ref/genomebiology.csl
output: 
  word_document:
    #reference_docx: C:/Users/Spiro Stilianoudakis/Documents/TAD_predictive_modeling/Manuscript/styles.doc/NIH_grant_style.docx
bibliography: C:/Users/Spiro Stilianoudakis/Documents/TAD_predictive_modeling/Manuscript/References/references.bib
---

Spiro C. Stilianoudakis^1^ (stilianoudasc@mymail.vcu.edu), Mikhail G. Dozmorov^1\*^ (mikhail.dozmorov@vcuhealth.org)

^1^ Dept. of Biostatistics, Virginia Commonwealth University, Richmond, VA, 23298, USA  
^\*^ To whom correspondence should be addressed: Virginia Commonwealth University, Richmond, VA, 23298, 804-827-2055, mikhail.dozmorov@vcuhealth.org

@@

**Methods.** We propose a novel data pre-processing pipeline that addresses those shortcomings. Lasso and Elastic Net were used for feature selection, coupled with five classifier methods. A number of classifier performance metrics were assessed, including the F1 measure and Matthew Correlation Coefficient (MCC), and area under the ROC curve (AUROC).

**Results.** Data preprocessing (log2-transformation and standardization) improves the performance of classification algorithms and allows for the ability to more accurately predict which genomic features are most associated with TAD boundaries. In contrast to overlap-based association measures, the distance between genomic elements was the most predictive measure. The random undersampling strategy addresses the class imbalance problem the most effectively in nearly all settings.

**Conclusions.** Current methods used to model the epigenomic features associated with TAD boundaries are insufficiently robust to handle properties of genomic data. Models applied to unprocessed data can have poor predictive performances. Focusing solely on standard performance assessment metrics, such as AUROC, can mask poor performance of the models; thus, the use of more balanced metrics, such as F1 and MCC, is warranted. Our model results in better performances and more accurate identification of important features associated with the formation of TAD boundaries.
